IIT CS595: Topics \& Applications in Programming Languages\\ {\large Homework 0: Rule induction, expressions, type safety}

### 2021-09-14 HW0/hw0.tex
complete task 2

\title{IIT CS595: Topics \& Applications in Programming Languages\\
  {\large Homework 0: Rule induction, expressions, type safety}}
\author{Your name here}

@@

  \[
  \begin{array}{l l}
    &
    \kwlet{x}{\kwn{2}}{x+(\kwlet{x}{\kwn{1}}{\kwlet{y}{x+\kwn{1}}{x+y}})}\\
    \step & ...\\
    \step & ...\\
  \end{array}
  \]
\end{task}

@@

>>>>

\title{IIT CS595: Topics \& Applications in Programming Languages\\
  {\large Homework 0: Rule induction, expressions, type safety}}
\author{Dustin Van Tate Testa}

@@

  \[
  \begin{array}{l l}
    &
    \kwlet{x}{\kwn{2}}{x+(\kwlet{x}{\kwn{1}}{\kwlet{y}{x+\kwn{1}}{x+y}})}\\
    \step & \kwn{2}+(\kwlet{x}{\kwn{1}}{\kwlet{y}{x+\kwn{1}}{x+y}})\\
    \step & \kwn{2}+(\kwlet{y}{\kwn{1}+\kwn{1}}{\kwn{1}+y}\\
    \step & \kwn{2}+(\kwlet{y}{\kwn{2}}{\kwn{1}+y}\\
    \step & \kwn{2}+(\kwn{1}+\kwn{2})\\
    \step & \kwn{2}+\kwn{3}\\
    \step & \kwn{5}\\
  \end{array}
  \]
\end{task}

@@

- Notice use of shorthand $\kwlen{\tree}$ to indicate the value of $n$ (black height) for tree \tree \\

Base Case: 
\begin{itemize}
    \item all cases when n = 1:
        \begin{itemize}
            \item {\tree = \leaf}: Rule~(RBT-1) confirms that when T is a single black node, n = 1.\\
            The proposition holds for this case as $\nodes{\leaf} = 1 \geq 2^{n-1} = 2^{1-1} = 1$.

            \item {\tree = \rnode{\leaf}{\leaf}}: rule~(RBT-3) confirms that when $\tree$ is a red node,
            n is the same as it's children, which when they're both leaves makes n = 1.\\
            The proposition holds for this case as follows:\\
            LHS: $\nodes{\rnode{\leaf}{\leaf}} = 1 + \nodes{\leaf} + \nodes{\leaf} = 1 + 1 + 1 = 3$\\
            RHS: $2^{n-1} = 2 ^ {1 - 1} = 2^0 = 1$\\
            giving $3 \geq 1$ 
        \end{itemize}
\end{itemize}

Inductive case assuming $\nodes{\tree} \geq 2^{n-1}$:
\begin{itemize}
    \item Case when tree's top element is a black node
        \[
        \begin{array}{l l l l}
            \nodes{\bnode{\tree_1}{\tree_2}} & \geq & 2^{\kwlen{\tree}-1} & \text{the proposition for this case}\\
            1 + \nodes{\tree_1} + \nodes{\tree_2} & \geq & 1 + 2 ^ {\kwlen{\tree_1} - 1} + 2 ^ {\kwlen{\tree_2} - 1} & \text{apply rule (RBT-2)}\\
            1 & \geq & 1  & \text{apply inductive hypothesis}\\
        \end{array}
        \]

    \item Case when tree's top element is a red node
        \[
        \begin{array}{l l l l}
            \nodes{\rnode{\tree_1}{\tree_2}} & \geq & 2^{\kwlen{\tree}-1} & \text{the proposition for this case}\\
            1 + \nodes{\tree_1} + \nodes{\tree_2} & \geq & 2 ^ {\kwlen{\tree_1} - 1} + 2 ^ {\kwlen{\tree_2} - 1} & \text{apply rule (RBT-3)}\\
            1 & \geq & 0  & \text{apply inductive hypothesis}\\
        \end{array}
        \]
\end{itemize}
